Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-08-12 Latex_files/Thesis/chapters/chapter 4/main.tex
updates to ch3

We repeated the same calculation but now while the dimer is in its 'inverted' orientation, instead of a single point where the optical force is minimised we 
see that there are instead two separate equilibrium positions, one above the focus
and one below the focus. In this particular example the two positions are far enough
apart that both can be considered as separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{50pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\caption{}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{30pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\caption{}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the centre of mass of 
		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
		the case where the dimer is in its' 'standard' orientation, where
		the dimer is trapped at $z = 1.71\ \mu m$. (c) is the case where 
		the dimer is in its' 'inverted' orientation, the dimer is trapped 
		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. 
		(b) and (d) are renders to visualise the dimer orientation are shown 
		below each plot. The black lines on each force-curve is a linear 
		fit with the slope being reported as the trap stiffness in the legend.}
		\label{fig:paradigmatic}
\end{figure}

We can see that both equilibrium positions have comparable axial trap 
stiffness ($\kappa_z$), however the difference in the transverse trap 
stiffness ($\kappa_x$) is far more noticeable. The same dimer was trapped
at each of the axial equilibrium positions and the transverse force was 
evaluated. While in all three cases the dimer can be trapped the linear
range where 
This highlights one of the challenges involved with studying 
asymmetric particles, even though its a simple enough process to trap them 
they maybe characterised very differently depending on their relative position 
and orientation towards the focus. This can have a significant impact on 
rheological studies - or attempting to probe any local property - as the 
variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
		where a positive force indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c). The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

@@

One aspect that has yet to be covered in depth with regards to
spherical aggregates of any construction is their interaction with
circularly polarised light. While most homogenous objects are trapped
no differently in different polarisations, dimers have been shown to
experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018}. 
Typically the spin density of an electric field cannot be reduced in 
homogenous medium due to the fact that the spin angular momentum is 
conserved locally. In which case the only way to transfer angular 
momentum to use an object that is anisotropic. This is seen most 
clearly with birefringent materials, due to fact that the crystal 
lattice itself is anisotropic, a circularly polarised beam will 
transfer an optical torque that is proportional to the difference in 
the refractive index of the particle's ordinary and extraordinary axis. 
The total optical torque is given as:
\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\Delta n$ is the difference in refractive indices of the 
particle, $\phi$ indicates the phase shift in the electric field
(for circular polarised light $\phi= \pi/4$), and $\theta$ is the angle
between the particle's main axis and the direction of the electric field.
The optical torque of non-birefringent spherical aggregates has no such 
relationship as the mechanism behind their rotation is not as clearly defined.

To visualise the relationship between the polarisation state of the 
trapping beam, and the rotation rate we simulated the motion of an 
optically trapped dimer in beams of varying polarisation ($NA=1.2$, 
$P=100\ mW$), and the dimer is composed of polystyrene ($n_p=1.59$, 
$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t 
=10^-5$) and at the end we looked at the orientational time series; 
the dimer's orientation is recorded as a quaternion which can be easily 
converted to a 3-dimensional rotation matrix. By considering only the 
transverse components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) 
of the rotation matrix and taking the Fourier transformation of their 
time series reveals the rotational frequency. The laser power is set to 
$100\ mW$ to avoid large rotational fluctuations and so that the Fourier 
series of the transverse components approximates $\delta(\omega_{rot}-f)$ 
- the Dirac delta function centred at the rotational frequency $\omega_{rot}$.

@@

A possible explanation of this phenomena was devised by by \cite{Yevick2017}, 
who found that highly focused Gaussian beams could produce second 
order effects in the Rayleigh regime resulting in a photo-kinetic 
force that results in orbital motion about the beam's central axis.
This was verified by \cite{Ruffner2012} who found that by trapping 
a silica bead and recording its position over the course of several 
minutes; while not immediately evident from the trajectory by measuring 
the probability flux of finding a particle in a given position 
revealed a helicity to the sphere's motion which matched the helicity 
of the trapping beam. This would hardly be considered similar to 
the behaviour shown by dimers; firstly because the probability flux 
is an indicator of the sphere's position not its orientation - the 
sphere can be thought to 'orbit' around the central beam axis with a 
radius of orbit on the order of $10^{-9}\ m$. Secondly, while 
\cite{Yevick2017} does compute a rotational frequency, for a single 
sphere this rotation rate is at most $0.01\ Hz$ whereas the lowest 
rotation rate recorded for a dimer ($a_{I}/a_{II}$) was around $2\ Hz$, 
a full two orders of magnitude greater. As such the rotation demonstrated
cannot be explained due to photo-kinetic forces.

This rotation was first noted by Vigilante and co-workers who only 
considered this behaviour for a symmetric dimer \cite{Vigilante2020}; 
to expand upon this work we looked at how the rotation rate changed
with the dimers' size ratio ($a_{I}/a_{II}$) and if the orientation
of the dimer plays any impact on the rotational frequency. By repeating
the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
a circularly polarised beam $\phi=90\circ$ it was found that not only is 
the rotation rate dependent on the size of the dimer, but also on its 
orientation and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while trapped in a
  	circularly polarised beam; a positive rotation rate indicates clockwise 
  	rotation, whereas a negative rotation rate indicates counter-clockwise 
  	rotation. The red line is for the case of a dimer in its 'standard' 
  	orientation. The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the beam. And 
  	lastly the green line is for the case when the dimer is in its 
  	'inverted' orientation, but when it is trapped below the focus of the 
  	beam.}
\end{figure}

@@

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity 
sized dimers there is the potential for non-vertical trapping 
orientations in which the dimer is located in a harmonic 
trap. When trapped by a circularly polarised light these 
dimers exhibit gyroscopic precession. As shown in fig~\ref{fig:gyro} 
the dimer's is rotating about it's long axis (as seen by 
the periodic behaviour of $U_x$ and $U_y$), while also precessing 
around the optical axis of the beam, resulting in a second 
order oscillation in $s_z$.
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
		off axis orientation with a circuarly polarised beam ($P= 100\ mW$). 
		The far left column depicts the dimer's centre of mass position with 
		time; middle two columns are the $x$, $y$, and $z$ components of the 
		vectors $u_x$ and $u_y$; last column depicts the components of the 
		vector $u_z$ which defines the dimer's orientation.}
	\label{fig:gyro}
\end{figure}

>>>>

We repeated the same calculation but now while the dimer is in its 'inverted' orientation, instead of a single point where the optical force is minimised we 
see that there are instead two separate equilibrium positions, one above the focus
and one below the focus. In this particular example the two positions are far enough
apart that both can be considered as separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.75\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{50pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\caption{}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.75\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{30pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\caption{}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the centre of mass of 
		the dimer (µm) for the case of a dimer of size ratio 2. (a) is 
		the case where the dimer is in its' 'standard' orientation, where
		the dimer is trapped at $z = 1.71\ \mu m$. (c) is the case where 
		the dimer is in its' 'inverted' orientation, the dimer is trapped 
		at two positions: $z = 1.20\ \mu m$ \& $z=-1.63\ \mu m$. 
		(b) and (d) are renders to visualise the dimer orientation are shown 
		below each plot. The black lines on each force-curve is a linear 
		fit with the slope being reported as the trap stiffness in the legend.}
		\label{fig:paradigmatic}
\end{figure}

We can see that both equilibrium positions have comparable axial trap 
stiffness ($\kappa_z$), however the difference in the transverse trap 
stiffness ($\kappa_x$) is far more noticeable. The same dimer was trapped
at each of the axial equilibrium positions and the transverse force was 
evaluated. While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable trap is far 
narrower in the 'standard' orientation compared to the 'inverted' cases. 
This highlights one of the challenges involved with studying 
asymmetric particles, even though its a simple enough process to trap them 
they maybe characterised very differently depending on their relative position 
and orientation towards the focus. This can have a significant impact on 
rheological studies - or attempting to probe any local property - as the 
variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's centre of mass spheres, 
		where a positive force indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation parameters are used here as in fig~\ref{fig:paradigmatic}(a) and (c). The blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

@@

One aspect that has yet to be covered in depth with regards to spherical 
aggregates of any construction is their interaction with circularly 
polarised light. For homogenous spheres the optical torque is regarded as
being negligible as the spin density cannot impart angular momentum while 
propagating in a homogenous medium. Dimers however, have been shown to
experience an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam.

First we wanted to confirm that the rotation seen in our simulations is
based on the polarisation of the trapping beam. To that end, we simulated 
the motion of an optically trapped dimer in beams of varying polarisation 
($NA=1.2$, $P=100\ mW$), the dimer is composed of polystyrene ($n_p=1.59$, 
$n_m=1.33$). Each simulation was run for $1$ second ($\Delta t =10^-5$) 
and at the end we looked at the orientational time series; the dimer's 
orientation is recorded as a quaternion which can be easily converted to 
a 3-dimensional rotation matrix. By considering only the transverse 
components ($U_{1,x}$, $U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation 
matrix and taking the Fourier transformation of their time series reveals 
the rotational frequency. The laser power is set to $100\ mW$ to avoid 
large thermal fluctuations and so that the Fourier series of the transverse 
components approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function 
centred at the rotational frequency $\omega_{rot}$.

@@

This shows us that these optical rotations are polarisation dependent 
and not merely an example of the dimer scattering light asymmetrically. 
The question then becomes, by what mechanism is the angular momentum of
the trapping beam being transferred to the dimer.

The rotation was first noted by Vigilante and co-workers who only 
considered this behaviour for a symmetric dimer \cite{Vigilante2020}.
In their work they attribute this to spin-curl effects, in which the 
curl of the spin density leads to a second order optical force that 
orbits around the beams central axis \cite{Yevick2017}. While several 
papers have demonstrated this phenomena \cite{Zhao2007,Zhao2009, Wang2010} 
it was only properly formalised by by \cite{Ruffner2012}. In which they 
showed that the seemingly random trajectory of a trapped sphere was biased
by the polarisation state of the trapping beam. While not immediately evident 
from the trajectory the helicity of the trapping beam was revealed by 
computing the particle's probability flux using.
\begin{align}
	j(r) = \frac{1}{N-1} \sum_{j=1}^{N-1}
	\frac{r_{j+1}-r_j}{\tau}\delta_{sigma_j}\left(r-\frac{r_{j+1}+r_j}{2}\right)
	\label{eq:prob_flux}
\end{align}

where $\delta_{\sigma_j}$ is the kernel of an adaptive density estimator
\cite{Silverman1986}. \eqref{eq:prob_flux} describes the direction a trapped
sphere is most likely move in given our statistical knowledge of the 
trajectories probability density function. A finite estimation of the density
function $p(r)$ is used in \cite{Ruffner2012}.
\begin{align}
	p(r) = \frac{1}{N}\sum_{j=1}^N \delta_{\sigma_j}(r-r_j)
	\label{eq:prob_density}
\end{align}

The probability flux reveals a biased motion in the trajectory of a single 
sphere (see Fig.~\ref{fig:Ruffner_Grier}). This biased motion results in a
slight orbital motion about the central axis of the trapping beam, the orbital
frequency is shown to be proportional to the polarisation state of the trapping
beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Ruffner_Grier_2012.pdf}
	\caption{Figure reused from \cite{Ruffner2012}. (a) shows how the momentum 
		density of a Gaussian beam is twisted while using circularly polarised 
		light. The top row (figures (b)-(d)) shows a 7 sphere cluster trapped 
		in a circularly polarised beam. Due to the clusters asymmetric 
		susceptibility to polarization the cluster rotates in the $x-y$ plane.   Whereas the bottom row (figures (e) - (g)) show the similar results for a single sphere. In this instance the sphere does not rotate but instead 
		orbits the beam axis. In both instances the motion is proportional to 
		the degree of polarisation (see figures (d) and (g)) but for the single
		sphere this motion is only revealed when using \eqref{eq:prob_flux} \& \eqref{eq:prob_density}. Reused with permission from author}
	\label{fig:Ruffner_Grier}
\end{figure}

While the results from \cite{Ruffner2012} suggest that the optical rotation 
seen in asymmetric dimers is attributed to the same spin-curl forces there 
are several discrepancies that cannot be explained purely by the spin-curl 
force. Firstly, there is the question of how the spin-curl force results in
an optical torque: Comparing figures~\ref{fig:Ruffner_Grier}(c) and (e) we 
can see that the behaviour of the 7 sphere cluster rotates about the central 
axis whereas the single sphere has an only slight orbital bias in its motion. 
The authors attributed this to the fact that the cluster is scattering light 
where the spin-curl force is more substantial due to it having a wider profile. 
This would suggest that particles' whose longer axis lies in the plane perpendicular
to the direction of propagation should experience any notable degree of 
torque. And indeed this appears to be the case for a number of experiments 
involving nano-dimers and ellipsoidal particles \cite{Ahn2018, Reimann2018}.

It does raise the question of how come our simulations show that despite the 
dimer being orientated so their long axis' align with the parallel to the 
direction of propagation they still readily rotate. You would not expect that
the addition of a single additional sphere should drastically adjust the torque
especially if said sphere is relatively small. However when we measured the 
optical torque of a single sphere and a dimer - $a_{I}/a_{II}=10$ - we found 
the exact opposite. In both cases we used the same trapping beam as used for
figure~\ref{fig:rotation_vs_pol} but with a circularly polarised beam. Both 
the sphere and dimer were rotated in the $x-z$ plane and the all three components 
of the optical torque were recorded.
\begin{figure}[h!]
	\centering	
	\includegraphics[width=\linewidth]{sphere_dimer_torque.png}
	\caption{Optical toque experienced by a dimer ($a_{I}/a_{II}=10$) and a 
	single isotropic sphere. Both were rotated in the $x-z$ plane and the angle 
	between $U_z$ and the beam axis gives the polar angle.The solid lines denote 
	the torque experienced by the dimer whereas the dashed lines represent the 
	torque experienced by the sphere.}
	\label{fig:sphere_dimer_torque}
\end{figure}

It should be noted that the difference in magnitude is not clearly shown in 
figure~\ref{fig:sphere_dimer_torque}. In truth, the optical torque acting on 
the dimer $10^{13}$ times greater than the torque applied to the single sphere. 
The torques about the $x$ and $y$ axis can be somewhat understood as the second 
sphere is being drawn back towards the centre of the trap by the gradient forces. 
The same cannot be said for the $z$component of the optical torque; while this 
effect could still be attributed to the spin-curl force, but it is clear that the internal scattering between the two spheres has some unintended effects.

 By repeating
the same kinds of simulation as used in \ref{fig:rotation_vs_pol} but for
a circularly polarised beam $\phi=90\circ$ it was found that not only is 
the rotation rate dependent on the size of the dimer, but also on its 
orientation and therefore their axial position.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{rotation_rate_vs_size.png}
  \caption{Rotation rate plotted against dimer size ratio while trapped in a
  	circularly polarised beam; a positive rotation rate indicates clockwise 
  	rotation, whereas a negative rotation rate indicates counter-clockwise 
  	rotation. The red line is for the case of a dimer in its 'standard' 
  	orientation. The blue line is for the case when the dimer is in its 
  	'inverted' orientation while trapped above the focus of the beam. And 
  	lastly the green line is for the case when the dimer is in its 
  	'inverted' orientation, but when it is trapped below the focus of the 
  	beam.}
\end{figure}

@@

\subsection{Gyroscopic Precession using asymmetric dimers}
As mentioned in section~\ref{sec:off-axis} for specificity sized dimers there 
is the potential for non-vertical trapping orientations in which the dimer is 
still stably trapped. When a circularly polarised beam is used the dimer exhibits gyroscopic precession. As shown in fig~\ref{fig:gyro} the dimer's trajectory
exhibits periodic rotation about its long axis $U_z$, precession motion as the
dimer spins around the central beam axis, and nutation motion as the dimer's
orientation rocks backwards and forwards. 
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{gyroscopic_precession.png}
	\caption{3 second trajectory of a dimer ($a_{I}/a_{II}=2$) trapped in an 
		off axis orientation with a circularly polarised beam ($P= 100\ mW$). 
		The far left column depicts the dimer's centre of mass position with 
		time; middle two columns are the $x$, $y$, and $z$ components of the 
		vectors $u_x$ and $u_y$; last column depicts the components of the 
		vector $u_z$ which defines the dimer's orientation.}
	\label{fig:gyro}
\end{figure}
